Pattern matching for\\ $(213,231)$-avoiding permutations

### 2015-06-14 article.tex
rewrite some corrolarry

\begin{corollary}
\label{corollary:minmaxelement}
Let $\pi \in S_n$.
Then
$\pi \in \AV_n(213,231)$ if and only if for $1 \leq i \leq n$,
$\pi[i]$ is the minimal or the maximal element of $\pi[i:]$.
\end{corollary}

\begin{corollary}
\label{corollary:max is ascent}
Let $\pi \in \in \AV_n(213,231)$ and $1 \leq i < n$.
Then,
(1) 
if $\pi[i]$ is an ascent element then $\pi[i]$ is the minimal element 
of $\pi[i:]$
and 
(2)
if if $\pi[i]$ is a descent element then $\pi[i]$ is the maximal element 
of $\pi[i:]$
\end{corollary}

@@

For the sake of readability, we now define the set
$\SET{\pmotif}{\ptext}{i}{j}$ as follows:
if $s \in \SET{\pmotif}{\ptext}{i}{j}$ then
$s$ is an occurrence of $\pmotif[\firsti{\pmotif}{i}:]$ in $\ptext[j:]$
and the element matching $\first{\pmotif}{i}$ is $\ptext[j]$.

\begin{lemma}
Let $\pmotif$ be a permutation where 
$\factor{\pmotif}{i}$ is an ascent (resp. descent) factor.
If $s \in \SET{\pmotif}{\ptext}{i}{j}$, such that the element matching $\first{\pmotif}{i-1}$ is minimal (resp. maximal), for all subsequence 
$s' \in \SET{\pmotif}{\ptext}{i}{j}$ and for all subsequence $t$ of $\ptext$, such that $t=t's'$, if $t$ is an occurrence of $\pmotif[\firsti{\pmotif}{i+1}:]$ then the subsequence $t's$ is an occurrence of $\pmotif[\firsti{\pmotif}{i+1}:]$.
\end{lemma}

@@

\begin{corollary}
Let $\pmotif$ be a permutation where $\factor{\pmotif}{i}$ is an ascent (resp. descent) factor.
If $s \in \SET{\pmotif}{\ptext}{i}{j}$, such that $s[\first{\pmotif}{i-1}]$ is minimal (resp. maximal). There is a match of $\pmotif$ in $\ptext$ with element $\first{\pmotif}{i}$ matched to $\ptext[j]$ if and only if there exists $t$ such that $ts$ is a match of $\pmotif$ in $\ptext$. 
\end{corollary}

@@

	\text{The maximal value of a subsequence} 
		& \text{If $\first{\sigma}{i}$ is  } \\
	\text{ which is an occurence of $\sigma[\firsti{\sigma}{i}:]$ in $\pi[j:]$} & 
		\text{ an ascent element } \\	
	\text{ and that minimize the maximal value} & \\
	\text{ and $\first{\sigma}{i}$ is matched with $\ptext[j]$} & \\

@@

	\text{The minimal value of a subsequence} 
		& \text{If $\first{\sigma}{i}$ is  } \\
	\text{ which is an occurence of $\sigma[\firsti{\sigma}{i}:]$ in $\pi[j:]$} & \text{ a descent element } \\	
	\text{ and that maximize the minimal value} & \\
	\text{ and $\first{\sigma}{i}$ is matched with $\ptext[j]$} & \\
\end{cases}
$$

@@

\noindent\textbf{BASE :}
\begin{align*}
\LM{\pmotif}{\ptext}{1}{j} 
&=
\begin{cases}
		\min_{j<j'}\{\infty\} \cup \{\ptext[j'] |\text{ $j'$ such that \indent } 
			& \text{If $\first{\sigma}{i}$ is}  \\
		\text{ $|\factor{\pmotif}{1}| \leq LIS(j,j',\ptext[j'])  $}\}&\text{ an ascent element } \\
		&\\
		\max_{j<j'}\{0\} \cup \{\ptext[j'] |\text{ $j'$ such that \indent } 
			& \text{If $\first{\sigma}{i}$ is}  \\
		\text{ $|\factor{\pmotif}{1}| \leq LDS(j,j',\ptext[j']) $}\}&\text{ a descent element }\\
\end{cases}
\\
\intertext{\textbf{STEP :}}
\LM{\pmotif}{\ptext}{i}{j} 
&=
\begin{cases}
	\min \{\infty\} \cup  \AF{\pmotif}{\ptext}{i}{j} &
	\text{If $\first{\sigma}{i}$ is an ascent element}\\
	\max \{0\} \cup  \DF{\pmotif}{\ptext}{i}{j} &
	\text{If $\first{\sigma}{i}$ is  a descent element}\\
\end{cases}
\end{align*}
where $\AF{\pmotif}{\ptext}{i}{j}$ and $\DF{\pmotif}{\ptext}{i}{j}$ are the sets 
of elements such that if  
$\ptext[k] \in \AF{\pmotif}{\ptext}{i}{j}$ or if $\ptext[k] \in \DF{\pmotif}{\ptext}{i}{j}$ 
then there exists a match of $\pmotif[\firsti{\pmotif}{i}:]$ in $\ptext[j:]$ and the element 
$\first{\pmotif}{i-1}$ is matched with element $\ptext[k]$. 
Given that one has the set $\SET{\pmotif}{\ptext}{i-1}{k}$, 
to decide if $\ptext[k] \in \AF{\pmotif}{\ptext}{i}{j}$, 
one needs to find an occurrence $t$ of $\factor{\pmotif}{i}$ in $\ptext[j:k-1]$ 
such there exists $s \in \SET{\pmotif}{\ptext}{i-1}{k}$ and $ts$ is a match of 
$\pmotif[\firsti{\pmotif}{i}:]$. 
Formally we define $\AF{\pmotif}{\ptext}{i}{j}$ and $\DF{\pmotif}{\ptext}{i}{j}$ as follows:
\begin{align*}
\AF{\pmotif}{\ptext}{i}{j}
&=
\text{$\{\ptext[j'] \;|\; j<j'$ and $\LM{\pmotif}{\ptext}{i-1}{j'} \neq 0$ and} 
\\
&\qquad 
\text{$|\factor{\pmotif}{i}| \leq LIS_{\ptext}(j,j'-1,\first{\pmotif}{i-1})\}$} 
\\
\DF{\pmotif}{\ptext}{i}{j}
&=
\text{$\{\ptext[j'] \;|\; j<j'$ and $\LM{\pmotif}{\ptext}{i-1}{j'} \neq \infty$ and}
\\
&\qquad
\text{$|\factor{\pmotif}{i}| \leq LDS_{\ptext}(j,j'-1,\first{\pmotif}{i-1})\}$}
\end{align*}
where $LIS_{\ptext}(i,j,k)$ (resp. $LDS_{\ptext}(i,j,k)$) is the longest increasing 
(resp. decreasing) sequence in $\ptext$ starting at $i$ and ending at $j$,
with every element of this sequence
smaller (resp. bigger) than $k$.
$LIS_{\ptext}$ and $LDS_{\ptext}$ can be computed in 
$O(|\ptext|^2\log(\log(|\ptext|)))$ (see \cite{Bespamyatnikh00enumeratinglongest}).

For the base case,
one is looking for a match for the first factor.
If the factor is an ascent (respectively descent) factor, then
one has to find an increasing (resp. decreasing) subsequence
in the text of same size or longest that
the size of the first factor.
To find the "safest" solution
one must assure that the right most element
of the sequence is minimal (resp. maximal).\\
For the induction, it follows the same idea
except that one must assure that
every element of the subsequence is larger (resp. smaller)
than the maximal (resp. minimal) element of the rest of the match which is given by the inductive problem of
the previous factor.

There exists a match of $\pmotif$ in $\ptext$ if and only if
there exists a $LM(n,i)\neq 0$ and $LM(n,i)\neq \infty$ for $1 \leq i \leq |\ptext|$,
with $n$ the number of factor in $\pmotif$.
Moreover the basic case can be computed in $O(n^2)$
and the induction on $O(kn^2)$ time and space.
\qed
\end{proof}

@@

\begin{lemma}
\label{lemma:ascentDescentAscent}
If $\overline{\sigma[i]\sigma[j]}$ (thus $\sigma[j]=\sigma[i]+1$) and 
if $\sigma[i]$ is an ascent (resp. decent) element and $\sigma[i]+1$ is an ascent 
(resp. decent) element then 
$i<j$ ($j>i$) and for every $k$, $i<k<j$ ($j>k>i$), $\pi[k]$ is a descent 
(resp. ascent) element.    
\end{lemma}

@@

	\end{cases}
	$$

	$\pmotif[i]$ can occur in $\ptext[j]$ because $\ptext[j] \in [\lb,\ub]$. $\pmotif[i:]$ occurs in $\ptext[j:]$, if and only if $\pmotif[i+1:]$ occurs in $\ptext[j+1:]$. $\pmotif[i+1:]$ occurs in $\ptext[j+1:]$ if and only if there is an occurrence of $\pmotif[i+1:]$ in $\ptext[j+1:]$ starting at element $k$, $j<k$, which correspond to know if $\bigcup_{j<k} \PM{\pbmotif}{\ptext}{*}{*}{i+1}{k}$ is true. Because $\pmotif[i]$ is an ascent element,  $\pmotif[i]$ is the minimal element of $\pmotif[i:]$, thus a match of $\pmotif[i+1:]$ can not use value smaller than $\ptext[j]$, so $\ptext[j]+1$ is a (new) lower bound. The upper bound is unchanged. So there is a occurrence of  $\pmotif[i:]$ starting at element $j$ in $\ptext[j:]$ if and only if $\bigcup_{j<k} \PM{\pbmotif}{\ptext}{\ptext[j]+1}{\ub}{i+1}{k}$.

@@

	$$
	\PM{\pbmotif}{\ptext}{\lb}{\ub}{i}{j}=
	\begin{cases}
			\bigcup_{j<k} \PM{\pbmotif}{\ptext}{\lb}{\ptext[j]-1}{i+1}{k}
				& \text{if $\pmotif[i]$ is not underlined } \\
				& \text{ and $\pmotif[i]$ is not overlined} \\
			\bigcup_{j<k} \PM{\pbmotif}{\ptext}{\lb}{\ptext[j]-1}{i+1}{k}
				& \text{if $\pmotif[i]$ is not underlined } \\
				& \text{ and $\overline{\pmotif[i](\pmotif[i]+1)}$ or ${\pmotif[i]}^\urcorner$}\\
				& \text{ and $\ptext[j]=\ub$} \\
			\PM{\pbmotif}{\ptext}{\lb}{\ptext[j]-1}{i+1}{j+1}
				& \text{if $\underline{\pmotif[i]\pmotif[i+1]}$ } \\
				& \text{ and $\overline{\pmotif[i](\pmotif[i]+1)}$ or ${\pmotif[i]}^\urcorner$}\\
				& \text{ and $\ptext[j]=\ub$} \\
			\PM{\pbmotif}{\ptext}{\lb}{\ptext[j]-1}{i+1}{j+1}
				& \text{if $\underline{\pmotif[i]\pmotif[i+1]}$ } \\
				& \text{ and $\pmotif[i]$ is not overlined} \\
			false & \text{otherwise}
	\end{cases}
	$$	
	We prove symmetrically as the last case that 
	$\pmotif[i:]$ occurs in $\ptext[j:]$
	with $\sigma[i]$ occurs in $\pi[j]$.
	 starting at element $j$ if and only if $\bigcup_{j<k} \PM{\pbmotif}{\ptext}{\lb}{\ptext[j]-1}{i+1}{k}$.

@@

Clearly if $\bigcup_{0<j} \PM{\pbmotif}{\ptext}{1}{n_\ptext}{1}{j}$ is true then $\sigma$ occurs in $\pi$. We now discuss about the position and value constraints.
\paragraph{Position Constraint.} There are 3 types of position constraints 
\begin{itemize}
	\item If $_\llcorner{\sigma[1]}$ then the left most element of $\sigma$  must occur in the left most element of $\pi$ ($\pmotif[1]$ occurs in $\ptext[1]$ on a occurrence of $\pmotif$ in $\ptext$). This constraint is assured by demanding that the occurrence starts at the left most element of $\ptext$ : if  $\PM{\pbmotif}{\ptext}{1}{n_\ptext}{1}{1}$ is true.

@@

	\begin{proposition}
	Let $\pi$ be a permutation.
	Computing the longest $(213,231)$-pattern that occurs in $\pi$
	can be done in $O(|\ptext|\log(\log(|\ptext|)))$ time and in $O(n)$ space.
	\end{proposition}

@@

\begin{proposition}
Given two permutations $\pi_1$ and $\pi_2$,
the longest common subsequence
avoiding (231,213) can be compute in $O(|\pi_1|^3|\pi_2|^3)$ time and space.
\end{proposition}

>>>>

\begin{corollary}
\label{corollary:minmaxelement}
$\pi \in \AV_n(213,231)$ if and only if for $1 \leq i \leq n$,
$\pi[i]$ is the minimal or the maximal element of $\pi[i:]$.
\end{corollary}

\begin{corollary}
\label{corollary:max is ascent}
Let $\pi \in \AV_n(213,231)$ and $1 \leq i < n$. Then,
(1) 
if $\pi[i]$ is an ascent element then $\pi[i]$ is the minimal element 
of $\pi[i:]$
and 
(2)
if if $\pi[i]$ is a descent element then $\pi[i]$ is the maximal element 
of $\pi[i:]$
\end{corollary}

@@

we now define the set
$\SET{\pmotif}{\ptext}{i}{j}$ as follows:
given a subsequence $s$ of $\ptext[j:]$,
$s \in \SET{\pmotif}{\ptext}{i}{j}$ if and only if
$s$ is an occurrence of $\pmotif[\firsti{\pmotif}{i}:]$ in $\ptext[j:]$
and the element$\first{\pmotif}{i}$ occurs in $\ptext[j]$.

\begin{lemma}
Let $\pmotif$ be a permutation 
with $\factor{\pmotif}{i}$ as an ascent (respectively descent) factor,
$s$ a subsequence such that, $s \in \SET{\pmotif}{\ptext}{i}{j}$ and $s[\first{\pmotif}{i-1}]$ is minimal (resp. maximal). 
For all subsequence 
$s' \in \SET{\pmotif}{\ptext}{i}{j}$ and for all subsequence $t$ of $\ptext$, such that $t=t's'$, if $t$ is an occurrence of $\pmotif[\firsti{\pmotif}{i+1}:]$ then the subsequence $t's$ is an occurrence of $\pmotif[\firsti{\pmotif}{i+1}:]$.
\end{lemma}

@@

\begin{corollary}
Let $\pmotif$ be a permutation 
with $\factor{\pmotif}{i}$ as an ascent (respectively descent) factor,
and let $s \in \SET{\pmotif}{\ptext}{i}{j}$, such that $s[\first{\pmotif}{i-1}]$ is minimal (resp. maximal). 
Those following statements are equivalent :
\begin{itemize}
	\item There exists an 
	occurrence of $\pmotif$ in $\ptext$ with element $\first{\pmotif}{i}$ occurs in $\ptext[j]$.
	\item There exists a subsequence $t$ of $\ptext[1:j-1]$  such that $ts$ is an occurrence of $\pmotif$ in $\ptext$ with element $\first{\pmotif}{i}$ occurs in $\ptext[j]$. 
\end{itemize}
\end{corollary}

@@

	\text{The maximal value of a subsequence} 
		& \text{If $\first{\sigma}{i}$ is  } \\
	\text{ which is an occurence of $\sigma[\firsti{\sigma}{i}:]$ in $\pi[j:]$} & 
		\text{ an ascent element } \\	
	\text{ and that minimize the maximal value} & \\
	\text{ and $\first{\sigma}{i}$ occurs in $\ptext[j]$} & \\

@@

	\text{The minimal value of a subsequence} 
		& \text{If $\first{\sigma}{i}$ is  } \\
	\text{ which is an occurence of $\sigma[\firsti{\sigma}{i}:]$ in $\pi[j:]$} & \text{ a descent element } \\	
	\text{ and that maximize the minimal value} & \\
	\text{ and $\first{\sigma}{i}$ occurs in $\ptext[j]$} & \\
\end{cases}
$$

@@

\noindent\textbf{BASE :}
\begin{align*}
\LM{\pmotif}{\ptext}{1}{j} 
&=
\begin{cases}
		\min_{j<j'}\{\infty\} \cup \{\ptext[j'] |\text{ $j'$ such that \indent } 
			& \text{If $\first{\sigma}{i}$ is}  \\
		\text{ $|\factor{\pmotif}{1}| \leq LIS(j,j',\ptext[j'])  $}\}&\text{ an ascent element } \\
		&\\
		\max_{j<j'}\{0\} \cup \{\ptext[j'] |\text{ $j'$ such that \indent } 
			& \text{If $\first{\sigma}{i}$ is}  \\
		\text{ $|\factor{\pmotif}{1}| \leq LDS(j,j',\ptext[j']) $}\}&\text{ a descent element }\\
\end{cases}
\\
\intertext{\textbf{STEP :}}
\LM{\pmotif}{\ptext}{i}{j} 
&=
\begin{cases}
	\min \{\infty\} \cup  \AF{\pmotif}{\ptext}{i}{j} &
	\text{If $\first{\sigma}{i}$ is an ascent element}\\
	\max \{0\} \cup  \DF{\pmotif}{\ptext}{i}{j} &
	\text{If $\first{\sigma}{i}$ is  a descent element}\\
\end{cases}
\end{align*}
where $\AF{\pmotif}{\ptext}{i}{j}$ and $\DF{\pmotif}{\ptext}{i}{j}$ are the sets 
of elements such that if  
$\ptext[k] \in \AF{\pmotif}{\ptext}{i}{j}$ or if $\ptext[k] \in \DF{\pmotif}{\ptext}{i}{j}$ 
then there exists an occurrence of $\pmotif[\firsti{\pmotif}{i}:]$ in $\ptext[j:]$ and the element 
$\first{\pmotif}{i-1}$ occurs in element $\ptext[k]$. 
Given that one has the set $\SET{\pmotif}{\ptext}{i-1}{k}$, 
to decide if $\ptext[k] \in \AF{\pmotif}{\ptext}{i}{j}$, 
one needs to find an occurrence $t$ of $\factor{\pmotif}{i}$ in $\ptext[j:k-1]$ 
such there exists $s \in \SET{\pmotif}{\ptext}{i-1}{k}$ and $ts$ is an occurrence of 
$\pmotif[\firsti{\pmotif}{i}:]$. 
Formally we define $\AF{\pmotif}{\ptext}{i}{j}$ and $\DF{\pmotif}{\ptext}{i}{j}$ as follows:
\begin{align*}
\AF{\pmotif}{\ptext}{i}{j}
&=
\text{$\{\ptext[j'] \;|\; j<j'$ and $\LM{\pmotif}{\ptext}{i-1}{j'} \neq 0$ and} 
\\
&\qquad 
\text{$|\factor{\pmotif}{i}| \leq LIS_{\ptext}(j,j'-1,\first{\pmotif}{i-1})\}$} 
\\
\DF{\pmotif}{\ptext}{i}{j}
&=
\text{$\{\ptext[j'] \;|\; j<j'$ and $\LM{\pmotif}{\ptext}{i-1}{j'} \neq \infty$ and}
\\
&\qquad
\text{$|\factor{\pmotif}{i}| \leq LDS_{\ptext}(j,j'-1,\first{\pmotif}{i-1})\}$}
\end{align*}
where $LIS_{\ptext}(i,j,k)$ (resp. $LDS_{\ptext}(i,j,k)$) is the longest increasing 
(resp. decreasing) sequence in $\ptext$ starting at $i$ and ending at $j$,
with every element of this sequence
smaller (resp. bigger) than $k$.
$LIS_{\ptext}$ and $LDS_{\ptext}$ can be computed in 
$O(|\ptext|^2\log(\log(|\ptext|)))$ (see \cite{Bespamyatnikh00enumeratinglongest}).

For the base case,
one is looking for an occurrence for the first factor.
If the factor is an ascent (respectively descent) factor, then
one has to find an increasing (resp. decreasing) subsequence
in the text of same size or longest that
the size of the first factor.
To find the "safest" solution
one must assure that the right most element
of the sequence is minimal (resp. maximal).\\
For the induction, it follows the same idea
except that one must assure that
every element of the subsequence is larger (resp. smaller)
than the maximal (resp. minimal) element of the rest of the occurrence which is given by the inductive problem of
the previous factor.

There exists a occurrence of $\pmotif$ in $\ptext$ if and only if
there exists a $LM(n,i)\neq 0$ and $LM(n,i)\neq \infty$ for $1 \leq i \leq |\ptext|$,
with $n$ the number of factor in $\pmotif$.
Moreover the basic case can be computed in $O(n^2)$
and the induction on $O(kn^2)$ time and space.
\qed
\end{proof}

@@

\begin{lemma}
\label{lemma:ascentDescentAscent}
If $\overline{\sigma[i]\sigma[j]}$ (thus $\sigma[j]=\sigma[i]+1$) and 
if $\sigma[i]$ is an ascent (respectively decent) element and $\sigma[i]+1$ is an ascent 
(resp. decent) element then 
$i<j$ (resp. $j>i$) and for every $k$, $i<k<j$ (resp. $j>k>i$), $\pi[k]$ is a descent 
(resp. ascent) element.    
\end{lemma}

@@

	\end{cases}
	$$	
	Remark that $\pmotif[i]$ can occur in $\ptext[j]$ because $\ptext[j] \in [\lb,\ub]$.
	Thus if $\pmotif[i+1:]$ occurs in $\ptext[j+1:]$ with every element in $[\pmotif[i]+1,\ub]$ then 
 	$\pmotif[i:]$ occurs in $\ptext[j:]$  	
 	which correspond to know if there exists $k$, $j<k$ such that
	$\PM{\pbmotif}{\ptext}{\ptext[j]+1}{\ub}{i+1}{k}$ is true.

@@

	$$
	\PM{\pbmotif}{\ptext}{\lb}{\ub}{i}{j}=
	\begin{cases}
			\bigcup_{j<k} \PM{\pbmotif}{\ptext}{\lb}{\ptext[j]-1}{i+1}{k}
				& \text{if $\pmotif[i]$ is not underlined } \\
				& \text{ and $\pmotif[i]$ is not overlined} \\
			\bigcup_{j<k} \PM{\pbmotif}{\ptext}{\lb}{\ptext[j]-1}{i+1}{k}
				& \text{if $\pmotif[i]$ is not underlined } \\
				& \text{ and $\overline{\pmotif[i](\pmotif[i]+1)}$ or ${\pmotif[i]}^\urcorner$}\\
				& \text{ and $\ptext[j]=\ub$} \\
			\PM{\pbmotif}{\ptext}{\lb}{\ptext[j]-1}{i+1}{j+1}
				& \text{if $\underline{\pmotif[i]\pmotif[i+1]}$ } \\
				& \text{ and $\overline{\pmotif[i](\pmotif[i]+1)}$ or ${\pmotif[i]}^\urcorner$}\\
				& \text{ and $\ptext[j]=\ub$} \\
			\PM{\pbmotif}{\ptext}{\lb}{\ptext[j]-1}{i+1}{j+1}
				& \text{if $\underline{\pmotif[i]\pmotif[i+1]}$ } \\
				& \text{ and $\pmotif[i]$ is not overlined} \\
			false & \text{otherwise}
	\end{cases}
	$$	
	We can show symmetrically as the last case that 
	$\pmotif[i:]$ occurs in $\ptext[j:]$
	with $\sigma[i]$ occurs in $\pi[j]$.
	 starting at element $j$ if and only if $\bigcup_{j<k} \PM{\pbmotif}{\ptext}{\lb}{\ptext[j]-1}{i+1}{k}$.

@@

Clearly if $\bigcup_{0<j} \PM{\pbmotif}{\ptext}{1}{n_\ptext}{1}{j}$ is true then $\sigma$ occurs in $\pi$. We now discuss about the position and value constraints.
\paragraph{Position Constraint.} There are 3 types of position constraints 
\begin{itemize}
	\item If $_\llcorner{\sigma[1]}$ then the left most element of $\sigma$  must occur in the left most element of $\pi$ ($\pmotif[1]$ occurs in $\ptext[1]$ on an occurrence of $\pmotif$ in $\ptext$). This constraint is assured by demanding that the occurrence starts at the left most element of $\ptext$ : if  $\PM{\pbmotif}{\ptext}{1}{n_\ptext}{1}{1}$ is true.

@@

	\begin{proposition}
	Let $\pi$ be a permutation. One can compute the
	the longest $(213,231)$-pattern that occurs in $\pi$
	in $O(|\ptext|\log(\log(|\ptext|)))$ time and in $O(n)$ space.
	\end{proposition}

@@

\begin{proposition}
Given two permutations $\pi_1$ and $\pi_2$,
one can compute 
the longest common subsequence
avoiding (231,213) in $O(|\pi_1|^3|\pi_2|^3)$ time and space.
\end{proposition}
